---
title: "clustering of fixations"
author: "Nils Wendel Heinrich"
date: '2024-09-23'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(dplyr)
library(arrow)
library(QuClu)

set.seed(36)

```

```{r eye_data, include=FALSE}

distant_fixations <- read_feather("data/Experiment3_DistantFixations.arrow")
close_fixations <- read_feather("data/Experiment3_CloseFixations.arrow")

data_fixations <- rbind(distant_fixations, close_fixations)

```
Fixations can be initiated with the spaceship either within foveal or peripheral vision. In literature, the threshold for an object to be in peripheral vision is 5 visual degrees. Meaning that if an object is 5° or further away from the center point of vision, it's in periphery. Can we recover this threshold for fixations that have a distance to a reference point (the spaceship)?

```{r clustering, include=FALSE}

# only have important columns in our data
ml3fix.clust <- data_fixations[c("distance_to_spaceship",
                                 "fixation_duration",
                                 "Dist_to_closest_obstacles",
                          	     "N_visible_obstacles",
                                 "input_noise",
                          	     "ID",
                          	     "level",
                                 "distant_fixation")] %>% na.omit()

# distant_fixation is =1, when the spaceship is further away than 5°. =0 otherwise.

cla <- kquantiles(ml3fix.clust[1], k=2, method="VS", B=50) # call cluster analysis

ml3fix.clust$cluster <- as.factor(cla$cl-1)		# add cluster assignment to data

# plotting
# by newly generated cluster
ggplot(ml3fix.clust, aes(x = distance_to_spaceship, colour = cluster)) +
  geom_density()

# looks like the algorithm roughly recovered the 5° threshold

# sanity check
max(ml3fix.clust[ml3fix.clust$cluster == 0, ]$distance_to_spaceship) # maximum distance of close fixations
min(ml3fix.clust[ml3fix.clust$cluster == 1, ]$distance_to_spaceship) # minimum distance of distant fixations

ml3fix.clust$disargeement <- as.numeric(ml3fix.clust$cluster) - as.numeric(ml3fix.clust$distant_fixation)
disargeements <- filter(ml3fix.clust, disargeement != 1)

# writing down data with clusters
write_feather(ml3fix.clust, "data/Experiment3_AllFixations_short.arrow")

```

# Result
The cluster algorithm draws the line at ~5.65°, visible by the maximum and minimum values for distance_to_spaceship in the individual clusters.
This does violate the 5° threshold established in literature somewhat. There are only 1507 disagreements.


```{r writing down data, include=FALSE}

# subsetting
distant_fixations <- ml3fix.clust[ml3fix.clust$cluster == 1, ]
close_fixations <- ml3fix.clust[ml3fix.clust$cluster == 0, ]

# julia arrow data
write_feather(distant_fixations, "data/Experiment3_DistantFixations_short.arrow")
write_feather(close_fixations, "data/Experiment3_CloseFixations_short.arrow")

```


